---
title: 'Election 2020: USA'
author: "BGL"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arm)
library(rvest)
library(maps)
library(ggthemes)
library(usmap)
library(socviz)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73",
    "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

**Better data than mine:**

https://github.com/favstats/USElection2020-EdisonResearch-Results/blob/main/data/latest/presidential.csv

Load and merge:

```{r}
url = "https://github.com/favstats/USElection2020-EdisonResearch-Results/raw/main/data/latest/presidential.csv"
results2020 = read.csv(url)
megafile = read.table("eday-covid.txt", header = TRUE)
library(dplyr)
results = left_join(results2020, megafile, by = "fips")
```

Add swings:

```{r}
results$Swing2020 = results$margin2020 - results$margin2016
results$Swing2016 = results$margin2016 - results$margin2012
```

### Some quick maps

Easy map:

```{r}
# Needs a fips variable
library(usmap)
plot_usmap(regions = "counties", include = "IN",
  data = results, values = "margin2020") +
  scale_fill_gradient2(low = "darkblue", mid = "darkorchid",
    high = "red", name = "Trump margin") +
  theme(legend.position = "right")
```

County map:

```{r}
us_states = map_data("state")
county_map$id = as.numeric(county_map$id)
# joining with election results 
county.election.df = left_join(county_map, results, by = c("id" = "fips"))
#county.election.df = drop_na(county.election.df)
```

```{r, fig.height=10, fig.width=16}
ggplot(county.election.df, aes(x = long, y = lat, fill = margin2020, group = group)) + geom_polygon(color = "gray90", size = 0.25) + theme_map() + labs(fill = "Trump margin") + scale_fill_gradient2(low = "blue", mid = "darkorchid", high = "red")
```

Nerf the color scale:

```{r}
nerf = county.election.df
nerf$margin2020[nerf$margin2020 > 25] = 25
nerf$margin2020[nerf$margin2020 < -25] = -25
```

```{r, fig.height=10, fig.width=16}
ggplot(nerf, aes(x = long, y = lat, fill = margin2020, group = group)) + geom_polygon(color = "gray90", size = 0.25) + theme_map() + labs(fill = "Trump margin") + scale_fill_gradient2(low = "blue", mid = "darkorchid", high = "red")
```

```{r, fig.height=10, fig.width=16}
nerfswing = county.election.df
nerfswing$Swing2020[nerfswing$Swing2020 > 10] = 10
nerfswing$Swing2020[nerfswing$Swing2020 < -10] = -10
ggplot(nerfswing, aes(x = long, y = lat, fill = Swing2020, group = group)) + geom_polygon(color = "gray90", size = 0.25) + theme_map() + scale_fill_gradient2(low = "blue", mid = "darkorchid", high = "red")
```




## Scatterplots

2020 swing vs. 2016 swing:

```{r}
results %>%
  ggplot(aes(Swing2016, Swing2020)) + 
  geom_smooth(method = "lm") +
  geom_point()
  #geom_point(data = filter(compresults, votes >= 50000))
```

2020 swing vs. college education:

```{r}
results %>%
  ggplot(aes(college, Swing2020)) + geom_smooth(se = FALSE) +
  geom_point(alpha = 0.1) +
  scale_x_log10() +
  xlab("Percent with a college degree (log scale)") +
  ylab("Swing (positive means Trump did better in 2020)") +
  ggtitle("Education polarization increased again in 2020") +
  labs(subtitle = "Trump improved in low education counties, did worse in high education counties")
```

Split by region:

```{r}
results %>%
  filter(census_region != "NA") %>%
  ggplot(aes(college, Swing2020)) + geom_smooth(se = FALSE) +
  geom_point(alpha = 0.1) +
  scale_x_log10() +
  xlab("Percent with a college degree (log scale)") +
  ylab("Swing (positive means Trump did better in 2020)") +
  ggtitle("Education polarization increased again in 2020") +
  labs(subtitle = "Trump improved in low education counties, did worse in high education counties") +
  facet_wrap(~census_region)
```

How does this compare to 2016?

```{r}
results %>%
  ggplot(aes(college, Swing2016)) + geom_smooth(se = FALSE) +
  geom_point(alpha = 0.3) +
  scale_x_log10()
```

Swing by Hispanic percentage:

```{r}
results$stateA = recode_factor(results$state.x, Florida = "Florida", Texas = "Texas", .default = "Everywhere else")
results %>%
  filter(stateA != "NA") %>%
  ggplot(aes(hisp_pct * 100, Swing2020)) + geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
#  scale_x_log10() +
  xlim(10, 100) +
  facet_wrap(~ stateA) +
  xlab("Hispanic percentage") +
  ylab("Swing (positive means Trump did better in 2020)") +
  ggtitle("Swing in counties with at least 10% Hispanic population")
```

Swing by COVID deaths per capita:

```{r}
results %>%
  ggplot(aes(deaths/popestimate2019, Swing2020)) +
  geom_point() +
  #geom_smooth(method = "lm") +
  geom_smooth(method = "gam", color = "orange") +
  scale_x_log10()
```



## Regression

Was 538 right?

```{r}
display(lm(margin2020 ~ margin2012 + margin2016, data = results))
```

What's the correlation of swing with (log) college education?

```{r}
cor(log(results$college), results$Swing2020, use = "pairwise")
```

```{r}
display(lm(Swing2020 ~ log(college), data = results))
```


Is swing related to race? (Note that since the percentages are small and the data is aggregated, this may be misleading.)

```{r}
display(lm(Swing2020 ~ black_pct + hisp_pct + asian_pct, data = results))
```

Does the Hispanic result hold in e.g. Illinois?

```{r}
results %>%
  filter(state.x == "Illinois") %>%
  ggplot(aes(hisp_pct, Swing2020)) + geom_text(aes(label = county))
```

Is there a relationship with COVID rate?

```{r}
display(lm(Swing2020 ~ log((cases+1)/popestimate2019), data = results))
```

Is there a relationship with COVID rate after accounting for college education?

```{r}
display(lm(Swing2020 ~ log((deaths+1)/popestimate2019) + log(college), data = results))
```




